Tabular Constraint Learning

### 2016-04-15 Paper/ECAI/tabular.tex
Completed table

\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
\newcommand{\format}[1]{\textit{#1}\xspace}
\newcommand{\generategroups}{\format{generateAssignments}}
\newcommand{\extractgroups}{\format{extractGroups}}
\newcommand{\extracttables}{\format{extractTables}}
\newcommand{\learnconstraints}{\format{learnConstraints}}
\newcommand{\findassignment}{\format{findSolutions}}
\newcommand{\postprocess}{\format{pruneRedundant}}
\newcommand{\constrainttorder}{\format{generalityOrder}}
\newcommand{\template}{\format{Constraint template}}

@@

We consider various sets of cells in tables:
\begin{itemize}
  \item {\bf Cells}: Each entry in the matrix/table is called  a \textit{cell}.
  We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$.
  \item {\bf Columns and rows}: To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
  \item
  A \textit{vector} is either a column or a row that is type-consistent.
  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
  \item
  A \textit{group} is a subrange of vectors with the same orientation in a table.
  We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
  We denote as the \textit{length} of a row (column) group $G$, written as \textit{length(G)}, the number of its columns (rows). We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
  \item
 A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$. If the subgroup $g$ must contain only one vector, we write $g \in G$.it defines what it means for a constraint to be well-formed
%We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
% A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
\end{itemize}
For notational convenience, we will refer to vectors with $V$ and to groups and subgroups with $G$.

@@

\subsection{Constraints}
A \template is a triple \textit{(\CName, \CSignature, \CFunction)}.
Let us elaborate on each of them. 
\begin{itemize}
\item
\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
\item
  \textit{\CSignature} is a set of constraints on the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName.
  For example, the signature can restrict the types of the groups, e.g., requiring them to be integers or constrain their sizes, e.g. the length of vectors in the arguments must be equal.
  The constraints act on properties or group meta-information, not on the actual group content. In ILP, it is known as bias, it defines what it means for a constraint be well-formed.
  \item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function. In ILP, it is known as background knowledge, which provides a definition to a predicate.
\end{itemize}

\paragraph{Example}
\sergey{need to use this below:}
  For example, a \template \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).

@@

While the traditional CSP problem is given a set of cells and a set of constraints,
and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
\\
{\bf Given }
\begin{itemize}
\item
a set of instantiated tables (so all cells have values) ${\cal T}$;
\item
a set of \template ${\cal C}$;
\item
a set of groups ${\cal G}$ for the tables ${\cal T}$;
\end{itemize}
\noindent
\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a \template}
{\bf Find}  the set $C\subseteq {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
for which
all $G_i \in {\cal G}$ and  3)  $c(G_1, ... , G_N)$ satisfies the signature of the constraint $c$.

@@

Let us now define what is a tabular solution for a \template in general. Let $C$ denote a \template, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.

\section{Problem Statement}
\sergey{should reflect all the changes, Luc wants to introduce constraint satisfaction into problem statement, maybe we should move the definition of a solution here? Like a solution is actually constraint satisfaction?}
In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}. Here we formalize the statement in terms of \template and group assignments as follows:

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
    & a constraint specificity DAG \dependencies \\
    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\
    & for each $c$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- \template's, \dependencies -- dependency graph
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constrainttorder(\constraints,\dependencies)$}
  \State $v_1, ..., v_n =$ variables of $c$
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S, \dependencies)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}
Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}. Let us elaborate on each step in detail.

@@

Finding candidate groups for a \template~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.

@@

\subsection{Subgroup satisfaction search}
$\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
Every \template~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.

@@

\subsection{Implementation note} Essentially for each \template both \generategroups and \findassignment are separate constraint satisfaction problems. Hence, it is natural to solve them using declarative solvers. Even though each is a separate constraint problem, they are variations of each other and can be modeled using similar techniques. It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem. Furthermore, declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.

@@

\newcommand{\numeric}{\format{numeric}}
\newcommand{\textual}{\format{textual}}
\newcommand{\integer}{\format{integer}}
\newcommand{\plength}{\format{length}}
\newcommand{\ptype}{\format{type}}
\newcommand{\ptable}{\format{table}}
\newcommand{\por}{\format{orientation}}
\newcommand{\nat}{\mathcal{N}}

\sergey{Table should take into account our new notation and balance between formality and intuition, heh}
\begin{table*}
  \centering
  \begin{tabularx}{\textwidth}{l X X}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
    \ecalldiff{v_x} & $\integer(v_x)$ or $\textual(v_x)$ & All elements in $v_x$ are different \\
    %X = Y & & \\
    \ecfkey{v_{fk}}{v_{pk}} & $\ptype(v_{fk})$ = $ptype(v_{pk})$,$\ptable(v_{fk}) \neq \ptable(v_{pk})$, $\ecalldiff{v_{pk}}$ & Every value in @fk also exist in @pk \\
    % TODO check out
    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
      & the variables in $\{@r, @1, @fk\}$ and in $\{@pk, @val\}$ all have the same \plength, \ptable and \por.
      & $@r$ is the obtained by multiplying $@1$ with $\eclookup{@fk}{@pk}{@val}$.
      \\
    \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
      & the variables in $\{@fk, @r\}$ and in $\{@pk, @val\}$ all have the same \plength, \ptable and \por; $@r$ and~$@val$ have the same type; and \ecfkey{@fk}{@pk}.
      & Lookup $@fk$ in $@pk$ and return the corresponding value in $@val$.
      \\
    \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}}
      & \textit{Same as lookup}
      & Lookup the first item in $@pk$ smaller than $@fk$ and return the corresponding value in $@val$.
      \\
    \ecperm{v_x}
      & $\numeric(@x)$, $\ecalldiff{@x}$
      & The values in $@x$ are a permutation of the numbers $1$ through $\plength(@x)$.
      \\
    \ecprod{v_r}{v_1}{v_2}
      & variables $\{@r, @1, @2\}$ are all $\numeric$ and have the same $\plength$
      & Every element in $@r$ is the product of the corresponding elements in~$@1$ and~$@2$.
      \\
    \ecproj{v_r}{\mathbf{v_x}}
      & $\{@r, @x\}$ all have the same $\plength$, $\por$, $\ptable$ and type; @x contains at least 2~vectors
      & At every position in $1$ through $\plength(@x)$ there is exactly one vector in $@x$ with a non-blank value at that position which is the same as the corresponding value in $@r$.
      \\
    \ecrank{v_r}{v_x}
      & $\integer(@r)$; $\numeric(@x)$; and $\plength(@r) = \plength(@x)$
      & The values in $@r$ represent the rank (from largest to smallest) of the values in ($@x$) - including ties
      \\
    \ectotal{v_r}{v_{pos}}{v_{neg}}
      & $\{@r, @pos, @neg\}$ all have the same $\plength$, which is at least $2$
      & The values in $@r$ are a running total, each value $@r[i] = @r[i - 1] + @pos[i] - @neg[i]$.
      \\
    \ecseries{v_x}
      & $\numeric(@x)$ and $\ecperm{@x}$
      & $@x[1] = 1$ and $@x[i] = @x[i - 1] + 1$.
      \\
    \ecsumc{v_r}{\mathbf{v_x}}
      &
      & Y is the result of summing each column in X \\
    \ecsumr{v_r}{\mathbf{v_x}} & & Y is the result of summing each row in X \\
    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
    \ecsumprod{v_r}{v_1}{v_2} & & \\

  \end{tabularx}
  \caption{\template's \sergey{Here a detailed explanation on what is going on, what is essential, what is not? etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}
}
  \label{table:constraints}
\end{table*}

>>>>

\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
\newcommand{\format}[1]{\textit{#1}\xspace}
\newcommand{\generategroups}{\format{generateAssignments}}
\newcommand{\extractgroups}{\format{extractGroups}}
\newcommand{\extracttables}{\format{extractTables}}
\newcommand{\learnconstraints}{\format{learnConstraints}}
\newcommand{\findassignment}{\format{findSolutions}}
\newcommand{\postprocess}{\format{pruneRedundant}}
\newcommand{\constrainttorder}{\format{generalityOrder}}

@@

We consider various sets of cells in tables:
\begin{itemize}
  \item {\bf Cells}: Each entry in the matrix/table is called  a \textit{cell}.
  We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$.
  \item {\bf Columns and rows}: To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
  \item
  A \textit{vector} is either a column or a row that is type-consistent.
  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
  \item
  A \textit{group} is a subrange of vectors with the same orientation in a table.
  We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
  We denote as the \textit{length} of a row (column) group $G$, written as \textit{length(G)}, the number of its columns (rows). We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
  \item
 A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$. If the subgroup $g$ must contain only one vector, we write $g \in G$.
%We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
% A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
\end{itemize}
For notational convenience, we will refer to vectors with $V$ and to groups and subgroups with $G$.

@@

\subsection{Constraints}
\sergey{write correspondence with ILP terminology}
A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
Let us elaborate on each of them. 
\begin{itemize}
\item
\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
\item
  \textit{\CSignature} is a set of constraints on the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName.
  For example, the signature can restrict the types of the groups, e.g., requiring them to be integers or constrain their sizes, e.g. the length of vectors in the arguments must be equal.
  The constraints act on properties or group meta-information, not on the actual group content. In ILP, it is known as mode declarations, when one needs to declare types of variables in the vocabulary. Also, it is analogous to standard linear algebra well-defined requirements, e.g., in order to multiply matrices $A$ and $B$, $A$ must have the size of $n \times k$ and $B$ the size $k \times m$.
\item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function. In ILP, it is known as bias, where we constrain ourselves to only particular subset of language in which we perform the search. 
\end{itemize}

\paragraph{Example}
\sergey{need to use this below:}
  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).

@@

While the traditional CSP problem is given a set of cells and a set of constraints,
and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
\\
{\bf Given }
\begin{itemize}
\item
a set of instantiated tables (so all cells have values) ${\cal T}$;
\item
a set of spreadsheet constraints ${\cal C}$;
\item
a set of groups ${\cal G}$ for the tables ${\cal T}$;
\end{itemize}
\noindent
\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a spreadsheet constraint}
{\bf Find}  the set $C\subseteq {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
for which
all $G_i \in {\cal G}$ and  3)  $c(G_1, ... , G_N)$ satisfies the signature of the constraint $c$.

@@

Let us now define what is a tabular solution for a spreadsheet constraint in general. Let $C$ denote a spreadsheet constraint, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.

\section{Problem Statement}
\sergey{should reflect all the changes, Luc wants to introduce constraint satisfaction into problem statement, maybe we should move the definition of a solution here? Like a solution is actually constraint satisfaction?}
In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}. Here we formalize the statement in terms of spreadsheet constraints and group assignments as follows:

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$\\
    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$,\\
    & a constraint specificity DAG \dependencies \\
    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\
    & for each $c$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Spreadsheet constraints, \dependencies -- dependency graph
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constrainttorder(\constraints,\dependencies)$}
  \State $v_1, ..., v_n =$ variables of $c$
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S, \dependencies)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}
Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}. Let us elaborate on each step in detail.

@@

Finding candidate groups for a spreadsheet constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.

@@

\subsection{Subgroup satisfaction search}
$\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
Every spreadsheet constraint~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.

@@

\subsection{Implementation note} Essentially for each spreadsheet constraint both \generategroups and \findassignment are separate constraint satisfaction problems. Hence, it is natural to solve them using declarative solvers. Even though each is a separate constraint problem, they are variations of each other and can be modeled using similar techniques. It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem. Furthermore, declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.

@@

\newcommand{\numeric}{\format{numeric}}
\newcommand{\textual}{\format{textual}}
\newcommand{\integer}{\format{integer}}
\newcommand{\discrete}{\format{discrete}}
\newcommand{\plength}{\format{length}}
\newcommand{\ptype}{\format{type}}
\newcommand{\ptable}{\format{table}}
\newcommand{\por}{\format{orientation}}
\newcommand{\prows}{\format{rows}}
\newcommand{\pcols}{\format{columns}}
\newcommand{\nat}{\mathcal{N}}

\sergey{Table should take into account our new notation and balance between formality and intuition, heh}
\begin{table*}
  \centering
  \begin{tabularx}{\textwidth}{l X X}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
    \ecalldiff{v_x} & $\discrete(v_x)$ & All values in $v_x$ are different: $v_x[i] \neq v_x[j]$ if $i \neq j$ \\
    %X = Y & & \\
    \ecperm{v_x}
      & $\numeric(v_{x})$, $\ecalldiff{v_{x}}$
      & The values in $v_{x}$ are a permutation of the numbers $1$ through $\plength(v_{x})$.
      \\
    \ecseries{v_x}
      & $\numeric(v_{x})$ and $\ecperm{v_{x}}$
      & $v_{x}[1] = 1$ and $v_{x}[i] = v_{x}[i - 1] + 1$.
      \\
    \ecfkey{v_{fk}}{v_{pk}} & $v_{fk}$ and~$v_{pk}$ are both $\discrete$; $\ptype(v_{fk}) = \ptype(v_{pk})$; $\ptable(v_{fk}) \neq \ptable(v_{pk})$; and $\ecalldiff{v_{pk}}$ & Every value in~$v_{fk}$ also exist in~$v_{pk}$ \\
    % TODO check out
    \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
      & $v_{fk}$ and $v_{pk}$ are both $\discrete$; variables $\{v_{fk}, v_{r}\}$ and $\{v_{pk}, v_{val}\}$ within the same set have the same \plength, \ptable and \por; $v_{r}$ and~$v_{val}$ have the same type; and \ecfkey{v_{fk}}{v_{pk}}.
      & $v_r[i]$ is the same value as looking up~$v_{fk}[i]$ in~$v_{pk}$  and returning the corresponding value in~$v_{val}$.
      \\
    \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}}
      & \textit{Same as lookup}
      & $v_r[i]$ is the same value as looking up the last item in~$v_{pk}$ smaller than~$v_{fk}[i]$ and returning the corresponding value in~$v_{val}$.
      \\
    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
      & Variables $\{v_{r}, v_{1}, v_{fk}\}$ are $\numeric$, variables $\{v_{pk}, v_{val}\}$ are $\discrete$ and within both sets all variables have the same \plength, \ptable and \por; also \ecfkey{v_{fk}}{v_{pk}}.
      & $v_{r}[i]$ is the obtained by multiplying $v_{1}[i]$ with $\eclookupf{v_{fk}}{v_{pk}}{v_{val}}{}[i]$.
      \\
    \ecprod{v_r}{v_1}{v_2}
      & Variables $\{v_{r}, v_{1}, v_{2}\}$ are all $\numeric$ and have the same $\plength$
      & $v_{r}[i] = v_{1}[i] \times v_{2}[i]$.
      \\
    \ecproj{v_r}{\mathbf{v_x}}
      & Variables $\{v_{r}, \mathbf{v_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{v_x}$ contains at least~2 vectors
      & At every position~$i$ in $1$ through $\plength(v_{r})$ there is exactly one vector~$v$ in $\mathbf{v_x}$ such that $v[i]$ is a non-blank value, then $v[i] = v_{r}[i]$.
      \\
    \ecrank{v_r}{v_x}
      & $\integer(v_{r})$; $\numeric(v_{x})$; and $\plength(v_{r}) = \plength(v_{x})$
      & The values in $v_{r}$ represent the rank (from largest to smallest) of the values in $v_{x}$ (including ties)
      \\
    \ectotal{v_r}{v_{pos}}{v_{neg}}
      & Variables $\{v_{r}, v_{pos}, v_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
      & The values in $v_{r}$ are a running total, each value $v_{r}[i] = v_{r}[i - 1] + v_{pos}[i] - v_{neg}[i]$.
      \\
    \ecsumc{v_r}{\mathbf{v_x}}
      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\prows(\mathbf{v_x}) \geq 2$; and $\pcols(\mathbf{v_x}) = \plength(v_r)$
      & Each value in $v_{r}$ is obtained by summing over the corresponding column in $\mathbf{v_x}$.
      \\
    \ecsumr{v_r}{\mathbf{v_x}}
      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\pcols(\mathbf{v_x}) \geq 2$; and $\prows(\mathbf{v_x}) = \plength(v_r)$
      & Each value in $v_{r}$ is obtained by summing over the corresponding row in $\mathbf{v_x}$.
      \\
    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}
      & $v_{fk}, v_{pk}$ are $\discrete$; $v_{r}, v_{val}$ are $\numeric$; within the sets $\{v_{val}, v_{fk}\}$ and $\{v_{pk}, v_{r}\}$ variables have the same $\plength$ and $\por$; $v_{fk}$ and $v_{val}$ have the same $\ptable$; $v_{fk}$ and $v_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{v_{pk}}
      & The value for $v_{r}[i]$ is obtained by summing all values $v_{val}[j]$ where $v_{fk}[j] = v_{pk}[i]$
      \\
    \ecsumprod{v_r}{v_1}{v_2}
      & Variables $\{v_r, v_1, v_2\}$ are all $\numeric$; $\plength(v_{1}) = \plength(v_{2}) \geq 2$; and $\prows(v_{r}) = \pcols(v_{r}) = 1$
      & $v_{r}[i] = \sum_{i = 1}^{\plength(v_{1})} v_{1}[i] \times v_{2}[i]$.
      \\

  \end{tabularx}
  \caption{Spreadsheet Constraints \sergey{Here a detailed explanation on what is going on, what is essential, what is not? etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}
}
  \label{table:constraints}
\end{table*}

\samuel{Satisfy signature for group assignment: find groups such that there exists a subgroup which satisfies the signature}
